A quantitative framework for investigating the reliability of network construction: supplemental information

### 2018-03-21 manuscript/CI_interactions.tex
removed comments from main text prior to upload

AC and DG wrote the code. AC performed statistical analyses. All authors contributed to writing and revising the manuscript. TR contributed dat. [[Anything missing?]]

@@

ecological networks; probabilistic interactions; Bayesian networks; sampling error; spatial variability; temporal variability

@@

\newpage
% Main text can be up to 7500 words, 10 fig., tables, boxes
\begin{spacing}{2.0}

@@

  Descriptions of ecological networks typically assume that the same interspecific interactions occur wherever and whenever a community is observed. This contrasts with the known stochasticity of ecological communities: community composition, species abundances, and link structure all vary in space and time. Moreover, finite sampling generates variation in the set of interactions actually observed. Here we develop the conceptual and analytical tools needed to capture uncertainty in the estimation of pairwise interactions. To define the problem, we identify the different contributions to the uncertainty of an interaction and its implications for the estimation of network properties. We then outline a framework to quantify the uncertainty around each interaction. We illustrate this framework using the most extensively sampled network to date. Finally, we offer both code and tangible recommendations for improving descriptions of ecological networks and sampling of rare interactions. Through these efforts, we demonstrate the utility of our approach and the importance of acknowledging the uncertainty inherent in network studies. Most importantly, we stress that networks are best thought of as systems constructed from random variables, the stochastic nature of which must be acknowledged for an accurate representation. Doing so will fundamentally change networks analyses, but should lead to greater realism.

  % Can be up to 200 words. This is 200 exactly.
  % Word count can be up to 7500, we're at about 6506 without the boxes, intro, references.

%DG: Eco Lett proposed the new format for methods papers, but I think they are nonetheless looking for conceptual contributions. We might have to tweak just a little bit the abstract to show that our proposition also help formalizing the idea that interactions should be described probabilistically. This is not only technical, it is fundamental as well.

% AC: Good idea - I've also changed the "In this study, we ..." paragraph to make it sound less like we just built some software and hopefully emphasise the conceptual side more.

% \linenumbers
\clearpage

@@

    Beyond ``true" variation in network structure, several researchers have pointed to the importance of sampling intensity for the assessment of network structure (e.g.,~\citealp{Martinez1999,Bluthgen2006,Bluthgen2007}). An assessment of the accumulation of interactions with increasing sampling effort suggests that it is even more challenging to document interactions than species~\citep{Poisot2012}. As a result, it has been proposed that interactions should be described probabilistically and network metrics computed accordingly~\citep{Poisot2016}. Early work in this vein includes food-web models using likelihood-based approaches~\citep{Allesina2008} or Gaussian~\citep{Williams2010} or binomial~\citep{Rohr2016} probability functions for each possible interaction. These models may include information about species' traits~\citep{Rohr2016} or may attempt to reproduce empirical network structures using a set of simple rules~\citep{Allesina2008,Williams2010}.
    % \citet{Eklof2013} also introduced variability  in the computation of network dimensions, allowing the absence of interactions within the n-dimension niche of every species. % AC: Does this really fit in now that this is in the introduction?

@@

   % [[Move this paragraph to intro?]] It is getting more and more common to represent interactions as probabilities in theoretical models of network structure. For example, the minimum potential model introduces gaps in the feeding niche of predators by sampling based on the likelihood of each interaction~\citep{Allesina2008}, while a Gaussian probability function is used to represent the niche by~\citet{Williams2010}. \citet{Eklof2013} also introduced variability  in the computation of network dimensions, allowing the absence of interactions within the n-dimension niche of every species. Empirical investigation of trait-matching constraints often consider the occurrence of interactions as binomial process (e.g.,~\citealp{Rohr2016}). There are also ways to perform analysis of probabilistic networks, with almost all metrics having a probabilistic version~\citep{Poisot2016}. But there is not yet consensus on what an interaction probability between a pair of species means, and this is definitely a problem when comes time to estimate the uncertainty of both pairwise interactions and entire networks.

    In this study, we formalize the description of interactions between species as probabilities and develop analytical tools to capture the uncertainty in the estimation of these interactions. We focus on binary interactions as a first step, but the framework could be expanded to deal with interaction frequencies and strength. To define the problem, we first identify the different contributions to the uncertainty of an interaction and discuss the implications of each source of uncertainty for the properties of ecological networks. Next, we develop an analytical framework to quantify the uncertainty around interactions in an empirical web. We illustrate this framework using the most extensively sampled network to date (Box 1).  Finally, we offer tangible recommendations for improved descriptors of ecological interactions. Through these efforts, we demonstrate both the utility of our approach and the importance of acknowledging the uncertainty inherent in network studies.

@@

     Even when an interaction is feasible, i.e. $L=1$, it may not occur at a given location or moment in time because of local constraints such as inclement weather or the lack of suitable habitat. We define the realisation of the interaction process with the variable $X$, given that the interaction is feasible, as a stochastic process with associated probability $P(X|L=1)=\chi$. This phenomenon of interaction contingencies is usually not considered in network studies, but there is a rich literature in community ecology about the contingencies of interactions. Phenological matching~\citep{MillerRushing2010,Gezon2016}, species preferences~\citep{Pires2011,Novak2015,Coux2016}, and fear effects of other species~\citep{Luttbeg2005,Wirsing2008} are just some of the factors contributing to variation in the frequency of interactions between a given pair of species. Although some of the factors leading to process uncertainty can be addressed in mesocosm studies of networks (e.g., environmental conditions can be held stable), process uncertainty is likely inevitable in the field. 
      % We imagine for instance that an interaction between a gall and a parasitoid might not be recorded at a location because it was the wrong time of the year and the gall had not yet formed (phenological constraints), a late frost killed the larvae (an abiotic environmental constraint), another parasitoid species competitively excluded the parasitoid species of interest (a biotic environmental constraint) or simply the parasitoid is too rare to interact with all of its viable hosts (an abundance constraint).

@@

    %DG: I do understand Tomas' point about the following, but I really appreciate it ! Keep in mind the suggestion to drop the following paragraph was made when there was much more material above. Perhaps we should revisit this suggestion, given the actual material that will follow. To me this is fairly important for the non-technical people. There are some important things, such as the "the more sample the more zeros will appear in the interaction matrix"

    %AE: I agree with Dom here, as I think the paragraphs below are good for getting a bit deeper into the width of the problem we are addressing. Could possibly be shortened if needed?

    %AC: I've added this section back in but shortened it slightly (mostly the first paragraph). We're ok on word-count with this included.

    % The basic information that we are collecting consists of matrices of interaction partners. Our count of interactions between two species reflects uncertainty with respect to whether two species are capable of interacting when co-occurring at a given site ($\lambda$), chance variation in whether the interaction is realised ($\chi$),  and uncertainty with respect to whether it is detected when realised ($\delta$). In this case, what can we do to pin down any of these different types of variation?

@@

\section*{A naive quantification of uncertainty}
  %AC: This section header was too long and running beyond a line... is the shortened version ok?
  % \section*{A naive quantification of interaction probability and uncertainty}

  %%AE: I think the section that is commented  below not should be placed here, but when reading the pdf without this part in it I really thoght that this was missing. One option is to move this part to the Introduction, where it would fit better.

   % [[Move this paragraph to intro?]] It is getting more and more common to represent interactions as probabilities in theoretical models of network structure. For example, the minimum potential model introduces gaps in the feeding niche of predators by sampling based on the likelihood of each interaction~\citep{Allesina2008}, while a Gaussian probability function is used to represent the niche by~\citet{Williams2010}. \citet{Eklof2013} also introduced variability  in the computation of network dimensions, allowing the absence of interactions within the n-dimension niche of every species. Empirical investigation of trait-matching constraints often consider the occurrence of interactions as binomial process (e.g.,~\citealp{Rohr2016}). There are also ways to perform analysis of probabilistic networks, with almost all metrics having a probabilistic version~\citep{Poisot2016}. But there is not yet consensus on what an interaction probability between a pair of species means, and this is definitely a problem when comes time to estimate the uncertainty of both pairwise interactions and entire networks.

@@

      %  The posterior distribution of $\theta$ could be computed with R with the following command:
      %
      %
      %  \vspace{12pt}
      %  \noindent\emph{dbeta(x = theta, shape1 = alpha+k, shape1 = beta+n-k)}
      %  \vspace{12pt}

@@

      It is common to preform analyses that require calculating higher-order network properties in interaction networks. The fact that the posterior distribution of $\lambda$ follows a beta distribution makes it straightforward to compute moments and other properties needed for this. 
    %%AE: Should we give examples of these higher-order properties?    
    % AC: How do you mean? Examples of averages/variances/modes or of "other properties" ?

@@

          The ecological network literature boasts a collection of networks for which connectance has been calculated and for which we can thus define the connectance distribution. Connectance is measured as $C = L/S^2$, where $L$ is the number of interactions and $S$ is the number of species. It measures the filling of an interaction matrix and thereby expresses the average probability that any two species interact with each other. If we know only the mean $\overline{C}$ and the variance $\sigma_C^2$ of the distribution of $C$,
          % [[from the literature or from our network? Has to be from the literature - if it was our network we'd know way more than just C]] 
          then the beta parameters could be computed as follows using the method of moments:

@@

          %  It is also possible to compute maximum likelihood estimates for a sample of the distribution of $C$.
          %  
          %   The following piece of R code provides an example: 
          %
          %  \vspace{12pt}
          %  \noindent\emph{
          %    library(MASS)\\
          %    pars = fitdistr(x = vecC, "beta", start = list(shape1 = 1, shape2 = 1))\$estimate
          %  }
          %  \vspace{12pt}
          %
          %
          %  Where $vecC$ is a vector of known connectances for a set of networks.

@@

    %It is fairly straightforward to compute most of network metrics when the different $\lambda$ of the adjacency matrix are known, without variance~\citep{Poisot2016}. Several of these metrics derive directly from quantitative indices of network structure which are equivalent to $\lambda$. The remainder, originally defined for binary networks, can be adjusted to account for interaction probabilities between zero and one. It is not as easy, however, to understand how the uncertainty in these estimated interaction probabilities influences network metrics. Computation of these metrics involves non-linear functions. Since Jensen's inequality states that the average of a non-linear function of a stochastic variable differs from the function of the average of that variable, any uncertainty in the values of $\lambda$ could bias both the mean and variance of a network metric. For instance, assume that among the pairs of species with few observed co-occurrences there are many pairs which truly interact. If we %begin with no observed links for these pairs, then correcting for these false absences will inflate the number of links and consequently influence metrics such as connectance, nestedness, and degree distribution. While we can, as described above, calculate these network metrics using fixed estimates for $\lambda$, no analytical solution currently exists for uncertain $\lambda$. Instead we must rely on computer simulations. 
    In the following section, we will provide an empirical example based on the well-sampled system of \emph{Salix}  plants, herbivorous gallers, and their natural enemies described by~\citet{Kopelke2017}; see Box~\ref{box1} or \emph{Appendix S2} for a description). Using this dataset, we will demonstrate the derivation of prior distributions for the \emph{Salix}-galler and galler-natural enemy components of these networks and the differences between these priors and posterior distributions which include all information available in this dataset~\citep{Kopelke2017}. Finally, we will calculate network properties using a suite of networks sampled from these posterior distributions and show how the uncertainty around interactions that have not been observed impact these metrics. %We find uncertainty influences both the average and the variance of network metrics, putting in perspective traditional network comparison studies where these metrics are assumed to be fixed entities.

    %DG: I dropped pieces of the above paragraph because it was bizarre to get back at the posterior distribution, and then to the computer simulations. This section is first and foremost an empirical example. But then, an introduction about the scaling up was missing, so I moved part of the paragraph below

    % AC : Maybe we should change the section title to highlight that this is an empirical example and covers the effect of uncertainty on network-level properties

@@

      For species where no co-occurrences were observed ($n=0$), we can calculate the estimates for the mean and variance of $\lambda_{ij}$ directly from the prior parameters following equations~\ref{mean} and~\ref{variance} (see \emph{Appendix S1} for R implementation). For the \emph{Salix}-galler network, the prior distribution  was: $\bar\lambda$=0.028, var($\lambda$)=8.60$\times$10$^{-5}$. The prior distribution for the galler-natural enemy network was: $\bar\lambda$=0.076, var($\lambda$)=0.008. The posterior interaction probabilities obtained based on the Zillis sub-network were much lower than those obtained based on~\citet[Data available from the Dryad Digital Repository: https://doi.org/10.5061/dryad.g7805]{Barbour2016}; this emphasizes the importance of using an appropriate study system when constructing a prior (\emph{Appendix S4}).

@@

    %DG: here is where I moved material from aboce

@@

    %DG: the different posterior connectance are the results of different samples of the posterior distribution, righ ? 
    %Yes

@@

    %    [[This is a little boring. Tomas has ideas for people with potentially better priors - may require another co-author. Also maybe could use one site as a prior like ``training data"?]]
    % DG: not so sure, I find it very useful because it simply illustrates the point. Nestedness will be a nice complement with lest trivial results
    %%AE: If we are doing a complement here (I agree it is a bit boring) I think option c) from Alyssas mail is the most straight forward option. 
    % AC: Added nestedness and replaced the Barbour data with training data from one site. Couldn't find any good published data.

@@

  % [Here we need to be very hands-on and remember the invitation letter form Ecol Letters:
  % As a contribution primarily to methodology, the manuscript will not be considered for ``Ideas \& Perspectives". (Please recall my message concerning your past presubmission inquiry.) However, Ecology Letters is entertaining the idea of a new section for ``Novel Methods". As we have not advertised this, I would ask you to keep that information in confidence. If acceptable to an Associate Editor and reviewers, your paper could be the first, or one of the first, contributions to this new section. Accordingly, the work would be subject to considerable scrutiny and I would urge you to make the paper as strong as you can. This would include demonstrating the need for this new method (i.e. outlining how this is different than other approaches currently in use and how they fail at some important task), documenting the robustness and accuracy of the approach, performing case study, and making robust high-quality code freely available.]

@@

  [[Anything to add here?]

@@

%%AE: I think Box 2 is nice to have in as a hands-on example.
\begin{floatbox}{}

@@

%DG: be careful : occurrence, not occurance on the figures
    \begin{figure}
      \caption{\textbf{A-B)} Most pairs of \emph{Salix} and gallers or gallers and natural enemies were never observed co-occurring despite the high levels of replication in our example dataset. For those pairs that were observed together at least once ($n_{ij}>0$), the number of observed co-occurrences was generally small (\textless10). Here we show a histogram of the number of pairs of species observed co-occurring at least once. 3986 \emph{Salix}-galler and 9794 galler-enemy pairs were never observed co-occurring: these pairs are omitted from the histogram. \textbf{C-D)} Most pairs of species that were observed at the same site were never observed interacting. Here we show a histogram of the number of observed interactions within pairs of co-occurring species. Species which co-occurred but never interacted are included in these histograms. \textbf{E-F)} Here we show, for each species pair, the number of observed interactions plotted against the number of observed co-occurrences. \emph{Salix}-galler pairs either are never observed interacting or interact almost every time they co-occur, while galler-enemy pairs had more variable frequencies of interaction. In panels E and F the red, dashed line indicates a 1:1 relationship between interactions and co-occurrences.}
      \label{histograms}
      \includegraphics*[width=.8\textwidth]{figures/Salix_Galler_histogram.eps}
      \end{figure}

%DG: I really appreciate this figure, it makes the point so clear.

% DG: it's bizarre to present here the outcome of the naive binomial formulation, while the reasoning is provided later. My solution is to move the following section on the binomial model along with the figure within a box. It's not central, but formalizes the problem. It would respect the order of appearance and better emphasize the bayes section.
% AC: I couldn't figure out how to put a figure inside a text box, so instead (following an e-mail suggestion) I moved the naive binomial outcome to the figure caption.

  \begin{figure}[h!]
    \caption{A simple example will illustrate the problem of imperfect detection of interactions. Assume that we want to infer the probability of an interaction between two species, $i$ and $j$. Now assume that in reality, interaction between $i$ and $j$ is completely impossible (i.e. the true $\lambda=0$) but the observer does not know this and seeks to estimate this interaction probability ($\lambda$). The number of observed interactions will follow a binomial distribution with number of interactions $k$ and number of observations $n$. Using this distribution, we can compute the credible interval of the estimated probability $\lambda$. Even assuming no added detection error in observing the incidence of the interaction, a single observation of species co-occurrence reveals very little regarding the probability of the interaction as the credible interval for a pair of species with one observation essentially spans from 0 to 1. Only with 35 observations will the upper limit of the credible interval be lowered to 0.1. Thus, adding more observations is certainly useful in controlling uncertainty, but the number of observations added needs to be very high. Here we show the upper bound (solid black line) of a 95\% Clopper-Pearson true credible interval for $\lambda$ when $k=0$ ($i$ and $j$ have not been observed interacting) for a variety of $n$ (observed co-occurrences of $i$ and $j$). Using a Bayesian approach with an informative prior can reduce the confidence interval about $\lambda$ for a given sample size. A threshold interaction probability of 0.1 is indicated by the dashed red line. }
    \label{upper_limits}
    \includegraphics*[width=.8\textwidth]{figures/upper_limit_DG.eps}
  \end{figure}

%DG :  I think we should add the MLE estimate along with the binomial confidence interval on the figure in order to better illustrate the impact of the prior (obviously only for one N). You may draw it with a line to contrast with polygons.
% Done, think it works for all N

%DG: otherwise the difference between the two interaction types is very interesting and useful to show the effect of the priors. Bravo !

      \begin{figure}[ht]
        \caption{Using prior distributions based on the \emph{Salix}-galler and galler-natural enemy networks sampled at a single site in~\citet{Kopelke2017}, we can calculate posterior distributions for the probability of interaction ($\lambda$) between two species that have not yet been observed interacting. Here we show posterior distributions for $\lambda$ in each network component ranging from the prior distribution ($n=0$ observed co-occurrence) to the distribution obtained when the pair of species has been observed co-occurring 100 times. The distribution narrows and approaches zero as the sample size increases. Likewise, the maximum likelihood estimator for the mean probability of interaction (diamonds at top of each panel) approaches zero and the 95\% credible interval (lines at top of each panel) narrows as sample size increases. \textbf{A)} The posterior distributions for $\lambda$ in the \emph{Salix}-galler component are narrower at low $n$ but shrink less with increased sampling than those for \textbf{B)} the distributions of $\lambda$ in the galler-natural enemy component.}
        \label{Salix_pdfs}
        \includegraphics*[width=.8\textwidth]{figures/Salix_Galler_pdfs_increasing_N_Zillis.eps}
        \end{figure}

      \begin{figure}[ht]
        \caption{The number of samples required to achieve a given level of confidence that an interaction probability $\lambda_{ij}$ is below a given threshold varies with both parameters. With a low threshold, our confidence that $\lambda_{ij}$ is below the threshold increases rapidly with repeated observation of co-occurrence without interaction. Here we show the cumulative density functions for threshold probabilities of 0.5 (solid line), 0.25 (dashed line), 0.1 (dash-dot line), and 0.05 (dotted line) as well as the points at which the cdf reaches 0.90 (orange square), 0.95 (red circle), and 0.975 (blue diamond) for each threshold value. The large ticks along the x-axis indicate the number of samples associated with each of these points. \textbf{A)} In the \emph{Salix}-galler network component, the 95\% credible interval for $\lambda_{ij}$ when $n$=0 was (0.013, 0.049). We can therefore be at least 95\% confident that $\lambda_{ij}$ is below thresholds of 0.1 or 0.05 without any observed co-occurrence of species $i$ and $j$. To be confident that $\lambda_{ij}$ is less than 0.01, however, would require more observed co-occurrences than there are sites in our dataset. \textbf{B)} In the galler-parasitoid network component, the 95\% credible interval for $\lambda_{ij}$ was substantially broader and many observed co-occurrences ($\approx$ 15-35) are required to be 95\% confident that $\lambda_{ij}$ is below thresholds of 0.1 or 0.05.}
        \label{Salix_cdfs}
        \includegraphics[width=.8\textwidth]{figures/Salix_Galler_samples_and_cdfs_Zillis.eps}

@@

    \begin{figure}[ht]
    \caption{Here we show the mean connectance, links per resource (\emph{Salix} in the \emph{Salix}-galler networks and gallers in the galler-natural enemy networks), links per consumer, and nestedness (NODF) for networks assembled using posterior distributions based on a single sub-network in the~\citet{Kopelke2017} dataset (Zillis). We created 100 ``posterior-sampling" networks and then, for each of these, created 100 ``detection-filter" networks by randomly sampling 50\%-99\% of the interactions included in the posterior-sampling network. This simulates imperfect detection of interactions in the field. Each point represents the mean network property (e.g., connectance) obtained from a set of 100 detection-filter networks, plotted against the value of the network property in the posterior-sampling network used to create the detection-filter networks. For each property and both network types, the posterior-sampling networks cover a relatively small range of network properties than the range covered by networks with varying detection probabilities. The value of each property decreases with the proportion of links included in the detection-filter networks.}
    \label{posterior_webs}    
    \includegraphics[width=.8\textwidth]{Figures/Salix_Galler_posterior_properties_Zillis.eps}
    \end{figure}

  % [[What if we have only one site? Can we use the number of observations of a pair of species interacting with any partners as the number of times they co-occur, and the number of interactions within the pair as k? This will be sort-of valid for insects but not for plants when researchers camp out by plant individuals... ]]

>>>>

DG designed the analytical approach. AC and DG wrote the code. AC performed statistical analyses. All authors contributed to writing and revising the manuscript.

@@

ecological networks; probabilistic interactions; Bayesian networks; sampling error; spatial variability; temporal variability; uncertainty

@@

\newpage

\begin{spacing}{2.0}

@@

  Descriptions of ecological networks typically assume that the same interspecific interactions occur each time a community is observed. This contrasts with the known stochasticity of ecological communities: community composition, species abundances, and link structure all vary in space and time. Moreover, finite sampling generates variation in the set of interactions actually observed. Here we develop the conceptual and analytical tools needed to capture uncertainty in the estimation of pairwise interactions. To define the problem, we identify the different contributions to the uncertainty of an interaction and its implications for the estimation of network properties. We then outline a framework to quantify the uncertainty around each interaction. We illustrate this framework using the most extensively sampled network to date. We found significant uncertainty in estimates for the probability of most pairwise interactions which we could, however, limit with informative priors. Through these efforts, we demonstrate the utility of our approach and the importance of acknowledging the uncertainty inherent in network studies. Most importantly, we stress that networks are best thought of as systems constructed from random variables, the stochastic nature of which must be acknowledged for an accurate representation. Doing so will fundamentally change networks analyses but should yield greater realism.
  % Can be up to 200 words. This is 200 exactly.

\linenumbers
\clearpage

@@

    Beyond ``true" variation in network structure, several researchers have pointed to the importance of sampling intensity for the assessment of network structure (e.g.,~\citealp{Martinez1999,Bluthgen2006,Bluthgen2007}). An assessment of the accumulation of interactions with increasing sampling effort suggests that it is even more challenging to document interactions than species~\citep{Poisot2012}. As a result, it has been proposed that interactions should be described probabilistically and network metrics computed accordingly~\citep{Poisot2016}. Early work in this vein includes food-web models using likelihood-based approaches~\citep{Allesina2008} or Gaussian~\citep{Williams2010} or binomial~\citep{Rohr2016} probability functions for each possible interaction. These models may include information about species' traits~\citep{Rohr2016} or may attempt to reproduce empirical network structures using a set of simple rules~\citep{Allesina2008,Williams2010}.

@@

    In this study, we formalise the description of interactions between species as probabilities and develop analytical tools to capture the uncertainty in the estimation of these interactions. We focus on binary interactions as a first step, but the framework could be expanded to deal with interaction frequencies and strength. To define the problem, we first identify the different contributions to the uncertainty of an interaction and discuss the implications of each source of uncertainty for the properties of ecological networks. Next, we develop an analytical framework to quantify the uncertainty around interactions in an empirical web. We illustrate this framework using the most extensively sampled network to date (Box 1).  Finally, we offer tangible recommendations for improved descriptors of ecological interactions. Through these efforts, we demonstrate both the utility of our approach and the importance of acknowledging the uncertainty inherent in network studies.

@@

     Even when an interaction is feasible, i.e. $L=1$, it may not occur at a given location or moment in time because of local constraints such as inclement weather or the lack of suitable habitat. We define the realisation of the interaction process with the variable $X$, given that the interaction is feasible, as a stochastic process with associated probability $P(X|L=1)=\chi$. This phenomenon of interaction contingencies is usually not considered in network studies, but there is a rich literature in community ecology about the contingencies of interactions. Phenological matching~\citep{MillerRushing2010,Gezon2016}, species preferences~\citep{Pires2011,Novak2015,Coux2016}, and fear effects of other species~\citep{Luttbeg2005,Wirsing2008} are just some of the factors contributing to variation in the frequency of interactions between a given pair of species. Although some of the factors leading to process uncertainty can be addressed in mesocosm studies of networks (e.g., environmental conditions can be held stable), process uncertainty is likely inevitable in the field.

@@

@@

\section*{A naive quantification of uncertainty}

@@

@@

      It is common to preform analyses that require calculating higher-order network properties in interaction networks. The fact that the posterior distribution of $\lambda$ follows a beta distribution makes it straightforward to compute moments and other properties needed for this.

@@

          The ecological network literature boasts a collection of networks for which connectance has been calculated and for which we can thus define the connectance distribution. Connectance is measured as $C = L/S^2$, where $L$ is the number of interactions and $S$ is the number of species. It measures the filling of an interaction matrix and thereby expresses the average probability that any two species interact with each other. If we know only the mean $\overline{C}$ and the variance $\sigma_C^2$ of the distribution of $C$,          then the beta parameters could be computed as follows using the method of moments:

@@

@@

    In the following section, we will provide an empirical example based on the well-sampled system of \emph{Salix}  plants, herbivorous gallers, and their natural enemies described by~\citet{Kopelke2017}; see Box~\ref{box1} or \emph{Appendix S2} for a description). Using this dataset, we will demonstrate the derivation of prior distributions for the \emph{Salix}-galler and galler-natural enemy components of these networks and the differences between these priors and posterior distributions which include all information available in this dataset~\citep{Kopelke2017}. Finally, we will calculate network properties using a suite of networks sampled from these posterior distributions and show how the uncertainty around interactions that have not been observed impact these metrics.

@@

      For species where no co-occurrences were observed ($n=0$), we can calculate the estimates for the mean and variance of $\lambda_{ij}$ directly from the prior parameters following equations~\ref{mean} and~\ref{variance} (see \emph{Appendix S1} for R implementation). For the \emph{Salix}-galler network, the prior distribution  was: $\bar\lambda$=0.028, var($\lambda$)=8.60$\times$10$^{-5}$. The prior distribution for the galler-natural enemy network was: $\bar\lambda$=0.076, var($\lambda$)=0.008. The posterior interaction probabilities obtained based on the Zillis sub-network were much lower than those obtained based on~\citet[Data available from the Dryad Digital Repository: https://doi.org/10.5061/dryad.g7805]{Barbour2016}; this emphasises the importance of using an appropriate study system when constructing a prior (\emph{Appendix S4}).

@@

@@

@@

@@

@@

  The authors thank Daniel Cartensen for fruitful discussion of the ideas in this manuscript. We also thank K\'{e}vin Cazalles for providing feedback on the manuscript.

@@

@@

    \begin{figure}
      \caption{\textbf{A-B)} Most pairs of \emph{Salix} and gallers or gallers and natural enemies were never observed co-occurring despite the high levels of replication in our example dataset. For those pairs that were observed together at least once ($n_{ij}>0$), the number of observed co-occurrences was generally small (\textless10). Here we show a histogram of the number of pairs of species observed co-occurring at least once. 3986 \emph{Salix}-galler and 9794 galler-enemy pairs were never observed co-occurring: these pairs are omitted from the histogram. \textbf{C-D)} Most pairs of species that were observed at the same site were never observed interacting. Here we show a histogram of the number of observed interactions within pairs of co-occurring species. Species which co-occurred but never interacted are included in these histograms. \textbf{E-F)} Here we show, for each species pair, the number of observed interactions plotted against the number of observed co-occurrences. \emph{Salix}-galler pairs either are never observed interacting or interact almost every time they co-occur, while galler-enemy pairs had more variable frequencies of interaction. In panels E and F the red, dashed line indicates a 1:1 relationship between interactions and co-occurrences.}
      \label{histograms}
      % \includegraphics*[width=.8\textwidth]{Salix_Galler_histogram.eps}
      \end{figure}

  \begin{figure}[h!]
    \caption{A simple example will illustrate the problem of imperfect detection of interactions. Assume that we want to infer the probability of an interaction between two species, $i$ and $j$. Now assume that in reality, interaction between $i$ and $j$ is completely impossible (i.e. the true $\lambda=0$) but the observer does not know this and seeks to estimate this interaction probability ($\lambda$). The number of observed interactions will follow a binomial distribution with number of interactions $k$ and number of observations $n$. Using this distribution, we can compute the credible interval of the estimated probability $\lambda$. Even assuming no added detection error in observing the incidence of the interaction, a single observation of species co-occurrence reveals very little regarding the probability of the interaction as the credible interval for a pair of species with one observation essentially spans from 0 to 1. Only with 35 observations will the upper limit of the credible interval be lowered to 0.1. Thus, adding more observations is certainly useful in controlling uncertainty, but the number of observations added needs to be very high. Here we show the upper bound (solid black line) of a 95\% Clopper-Pearson true credible interval for $\lambda$ when $k=0$ ($i$ and $j$ have not been observed interacting) for a variety of $n$ (observed co-occurrences of $i$ and $j$). Using a Bayesian approach with an informative prior can reduce the confidence interval about $\lambda$ for a given sample size. A threshold interaction probability of 0.1 is indicated by the dashed red line. }
    \label{upper_limits}
    % \includegraphics*[width=.8\textwidth]{figures/upper_limit_DG.eps}
  \end{figure}

      \begin{figure}[ht]
        \caption{Using prior distributions based on the \emph{Salix}-galler and galler-natural enemy networks sampled at a single site in~\citet{Kopelke2017}, we can calculate posterior distributions for the probability of interaction ($\lambda$) between two species that have not yet been observed interacting. Here we show posterior distributions for $\lambda$ in each network component ranging from the prior distribution ($n=0$ observed co-occurrence) to the distribution obtained when the pair of species has been observed co-occurring 100 times. The distribution narrows and approaches zero as the sample size increases. Likewise, the maximum likelihood estimator for the mean probability of interaction (diamonds at top of each panel) approaches zero and the 95\% credible interval (lines at top of each panel) narrows as sample size increases. \textbf{A)} The posterior distributions for $\lambda$ in the \emph{Salix}-galler component are narrower at low $n$ but shrink less with increased sampling than those for \textbf{B)} the distributions of $\lambda$ in the galler-natural enemy component.}
        \label{Salix_pdfs}
        % \includegraphics*[width=.8\textwidth]{figures/Salix_Galler_pdfs_increasing_N_Zillis.eps}
        \end{figure}

      \begin{figure}[ht]
        \caption{The number of samples required to achieve a given level of confidence that an interaction probability $\lambda_{ij}$ is below a given threshold varies with both parameters. With a low threshold, our confidence that $\lambda_{ij}$ is below the threshold increases rapidly with repeated observation of co-occurrence without interaction. Here we show the cumulative density functions for threshold probabilities of 0.5 (solid line), 0.25 (dashed line), 0.1 (dash-dot line), and 0.05 (dotted line) as well as the points at which the cdf reaches 0.90 (orange square), 0.95 (red circle), and 0.975 (blue diamond) for each threshold value. The large ticks along the x-axis indicate the number of samples associated with each of these points. \textbf{A)} In the \emph{Salix}-galler network component, the 95\% credible interval for $\lambda_{ij}$ when $n$=0 was (0.013, 0.049). We can therefore be at least 95\% confident that $\lambda_{ij}$ is below thresholds of 0.1 or 0.05 without any observed co-occurrence of species $i$ and $j$. To be confident that $\lambda_{ij}$ is less than 0.01, however, would require more observed co-occurrences than there are sites in our dataset. \textbf{B)} In the galler-parasitoid network component, the 95\% credible interval for $\lambda_{ij}$ was substantially broader and many observed co-occurrences ($\approx$ 15-35) are required to be 95\% confident that $\lambda_{ij}$ is below thresholds of 0.1 or 0.05.}
        \label{Salix_cdfs}
        % \includegraphics[width=.8\textwidth]{figures/Salix_Galler_samples_and_cdfs_Zillis.eps}

@@

    \begin{figure}[ht]
    \caption{Here we show the mean connectance, links per resource (\emph{Salix} in the \emph{Salix}-galler networks and gallers in the galler-natural enemy networks), links per consumer, and nestedness (NODF) for networks assembled using posterior distributions based on a single sub-network in the~\citet{Kopelke2017} dataset (Zillis). We created 100 ``posterior-sampling" networks and then, for each of these, created 100 ``detection-filter" networks by randomly sampling 50\%-99\% of the interactions included in the posterior-sampling network. This simulates imperfect detection of interactions in the field. Each point represents the mean network property (e.g., connectance) obtained from a set of 100 detection-filter networks, plotted against the value of the network property in the posterior-sampling network used to create the detection-filter networks. For each property and both network types, the posterior-sampling networks cover a relatively small range of network properties than the range covered by networks with varying detection probabilities. The value of each property decreases with the proportion of links included in the detection-filter networks.}
    \label{posterior_webs}    
    % \includegraphics[width=.8\textwidth]{Figures/Salix_Galler_posterior_properties_Zillis.eps}
    \end{figure}

### 2018-03-21 manuscript/coverletter.tex
removed comments from main text prior to upload

\position{Postdoctoral fellow}
\department{Department of Physics, Chemistry, and Biology (IFM)}
\location{581 83 Link\"{o}ping, Sweden}
\cityzip{581 83 Link\"{o}ping, Sweden}
\telephone{}
\fax{}
\email{alyssa.cirtwill@liu.se}
\url{http://cirtwill.github.io}
\name{Alyssa R. Cirtwill}

@@

    My co-authors and I are pleased to submit the attached manuscript \emph{A quantitative framework for investigating the reliability of network construction} for consideration as a Novel Methods paper in \emph{Ecology Letters} as per our previous correspondence with Senior Editor John Drake. As instructed, we have submitted the manuscript under the Ideas \& Perspectives mechanism (and following the associated formatting guidelines).

@@

    We appreciate the chance to submit one of the first Novel Methods papers at \emph{Ecology Letters}. We are confident that the fundamental conceptual shift implied by our approach, accompanied by clear examples and easy-to-use scripts, can achieve the high standard you are looking for. We hope that you agree and eagerly await your reply.

@@

>>>>

\position{Postdoctoral fellow}
\department{Department of Physics, Chemistry, and Biology (IFM)}
\location{581 83 Link\"{o}ping, Sweden}
\cityzip{581 83 Link\"{o}ping, Sweden}
\telephone{}
\fax{}
\email{alyssa.cirtwill@liu.se}
\url{http://cirtwill.github.io}
\name{Dr. Alyssa R. Cirtwill}

@@

    My co-authors and I are pleased to submit the attached manuscript \emph{A quantitative framework for investigating the reliability of network construction} for consideration as a Novel Methods paper in \textbf{Ecology Letters} as per our previous correspondence with Senior Editor John Drake. As instructed, we have submitted the manuscript under the Ideas \& Perspectives mechanism (and following the associated formatting guidelines).

@@

    We appreciate the chance to submit one of the first Novel Methods papers at \textbf{Ecology Letters}. We are confident that the fundamental conceptual shift implied by our approach, accompanied by clear examples and easy-to-use scripts, can achieve the high standard you are looking for. We hope that you agree and eagerly await your reply.

@@

% Pedro Jordano pedroj@me.com Estacion Biologica de Donana CSIC
% Phillip Staniczenko pstaniczenko@sesync.org National Socio-Environmental Synthesis Center (SESYNC)
